---
title: "Module1_Quiz"
author: "mindan"
date: "2021/10/3"
output: pdf_document
---

```{r setup, eval=TRUE}
knitr::opts_chunk$set(cache = TRUE)
```

```{r }
rm(list=ls())
```

3.Create a `summarizedExperiment` object with the following code

```{r }
rm(list=ls())
library(Biobase)
library(GenomicRanges)
library(SummarizedExperiment)
data(sample.ExpressionSet, package = "Biobase")
se = SummarizedExperiment(sample.ExpressionSet)
assays(se)
colData(se)
rowData(se)
rowRanges(se)

```
## Load the Bottomly and the Bodymap data sets with the following code
```{r }
con =url("http://bowtie-bio.sourceforge.net/recount/ExpressionSets/bottomly_eset.RData")
load(file=con)
close(con)
bot = bottomly.eset
pdata_bot=pData(bot)
pdata_bot

con =url("http://bowtie-bio.sourceforge.net/recount/ExpressionSets/bodymap_eset.RData")
load(file=con)
close(con)
bm = bodymap.eset
pdata_bm=pData(bm)
pdata_bm
edata_bm = exprs(bm)
```
5. Just considering the phenotype data what are some reasons that the Bottomly data set is likely a better experimental design than the Bodymap data? Imagine the question of interest in the Bottomly data is to compare strains and in the Bodymap data it is to compare tissues.

```{r}
library(plotrix)
pie3D(pdata_bm$num.tech.reps,labels=pdata_bm$tissue.type)
```
6. What are some reasons why this plot is not useful for comparing the number of technical replicates by tissue (you may need to install the plotrix package).

7.Which of the following code chunks will make a heatmap of the 500 most highly expressed genes (as defined by total count), without re-ordering due to clustering? Are the highly expressed samples next to each other in sample order?
```{r}
row_sums = rowSums(edata_bm)
index = which(rank(-row_sums) < 500 )
heatmap(edata_bm[index,],Rowv=NA,Colv=NA)

```
8.Make an MA-plot of the first sample versus the second sample using the log2 transform (hint: you may have to add 1 first) and the `rlog` transform from the DESeq2 package. How are the two MA-plots different? Which kind of genes appear most different in each plot?

```{r}
mm = log2(edata_bm[,1]+1) - log2(edata_bm[,2]+1)
aa = log2(edata_bm[,1]+1) + log2(edata_bm[,2]+1)
plot(aa,mm,col=2)

library(DESeq2)
edata1 <- rlog(edata_bm)
mm2 = edata1[,1] - edata1[,2]
aa2 = edata1[,1] + edata1[,2]
plot(aa2,mm2,col=2)

```


## Load the Montgomery and Pickrell eSet
```{r}
con =url("http://bowtie-bio.sourceforge.net/recount/ExpressionSets/montpick_eset.RData")
load(file=con)
close(con)
mp = montpick.eset
pdata=pData(mp)
edata=as.data.frame(exprs(mp))
fdata = fData(mp)
```
9. Cluster the data in three ways:

- With no changes to the data

- After filtering all genes with` rowMeans` less than 100

- After taking the `log2` transform of the data without filtering

Color the samples by which study they came from (Hint: consider using the function `myplclust.R` in the package `rafalib` available from CRAN and looking at the argument `lab.col`.)

How do the methods compare in terms of how well they cluster the data by study? Why do you think that is?
```{r}
library(rafalib)
colramp = colorRampPalette(c(3,"white",2))(9)
# cluster With no changes to the data
dist1 = dist(t(edata))
heatmap(as.matrix(dist1),col=colramp,Colv=NA,Rowv=NA)
hclust1 = hclust(dist1)
myplclust(hclust1, labels = hclust1$labels, lab.col = rep(1, length(hclust1$labels)), hang = 0.1)

# After filtering all genes with rowMeans less than 100
edata2 = edata[rowMeans(edata) > 100,]
dist1 = dist(t(edata2))
heatmap(as.matrix(dist1),col=colramp,Colv=NA,Rowv=NA)
hclust1 = hclust(dist1)
plot(hclust1)

# After taking the log2 transform of the data without filtering
edata3 = log2(edata + 1)
dist1 = dist(t(edata3))
heatmap(as.matrix(dist1),col=colramp,Colv=NA,Rowv=NA)
hclust1 = hclust(dist1)
plot(hclust1)
```

10.Cluster the samples using k-means clustering after applying the `log2` transform (be sure to add 1). Set a seed for reproducible results (use `set.seed(1235)`). If you choose two clusters, do you get the same two clusters as you get if you use the `cutree` function to cluster the samples into two groups? Which cluster matches most closely to the study labels?

```{r}
edata3 = log2(edata + 1)
dist1 = dist(t(edata3))
hclust1 = hclust(dist1)
plot(hclust1)
cut <- cutree(hclust1, k = 2)
table(cut)

set.seed(1235)
kmeans1 = kmeans(t(edata3),centers=2)
names(kmeans1)
table(kmeans1$cluster)
heatmap(as.matrix(edata)[order(kmeans1$cluster),],col=colramp,Rowv=NA)
all.equal(kmeans1$cluster,cut)

#labels_colors(dend) = c(rep(1,60),rep(2,69))
#plot(dend)
```

```{r session_info}
devtools::session_info()
```